First and foremost, I would like to thank my research supervisors Ph.D Edgardo Solano Carillo and Prof. Henrik Lipskoch very much for their excellent support during my time of preparing and writing this thesis. Their direct feedback and the numerous discussions we had steered me in the right direction whenever I needed it during the six month of planning, experimenting and putting it all together. At no time did I feel left alone or overwhelmed, because I knew that they were always there to give me competent advice.
\par
Besides, I would also like to thank the Institute for the Protection of Maritime Infrastructures from the German Aerospace Center (DLR) for giving me the opportunity to write my master thesis at the institute. I was able to experience the world of the research community first-hand while meeting and having chats with wonderful and highly skilled colleagues. The institute also provided the necessary historical AIS records from the coast of Bremerhaven.
\par
I would also like to acknowledge all the proof readers, the people who commented and those who had to endure my long monologues when thinking out loud, especially my twin brother Dirk, my girlfriend Lena and my close friends Lukas and Simon. Amazingly, they all studied and work in the field of IT and therefore gave valuable criticism and continuous encouragement throughout my years of study and the progress of writing this thesis.
\par
Once again, thank you to all who took part in the creation of this work.
\par
Author
\par
Jan Löwenstrom